### 2020-04-30 Final_Draft.Rmd
changes to appendix and tables

---
title: "Investigating the Underprovision of Public Goods: Does Segregation Matter?"
author: "Maria Burzillo"
date: "5/8/2020"
output: 
  pdf_document:
    fig_caption: yes
bibliography: final_proj_bib.bib
#biblio-style: "apalike"
link_citations: true
keep_tex: yes
header_includes: \usepackage{float}

---

# Abstract

@trounstine_segregation_2016 suggests that high levels of residential segregation are associated with increased political polarization and decreased public spending. In this analysis, I was able to successfully replicate @trounstine_segregation_2016’s main results. I then attempted to better deal with the large amounts of missing data in the datasets used in the original analysis by multiply imputing missing values using the mice() package and then by re-running the original models using the multiply imputed datasets. My results suggest that segregation is associated with increased racial political polarization, although maybe not as strongly as Trounstine (2016) originally suggested. Furthermore, I find that Trounstine (2016)'s conclusion that increases in segregation are associated with decreases in public spending holds for large cities, but that diversity is a better explanatory factor for small cities.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.pos = "!H")
knitr::write_bib(c("knitr", "stringr"), "bib.bib", width = 60)
library(haven)
library(gtable)
library(stargazer)
library(tidyverse)
library(gt)
library(gtsummary)
library(knitr)
library(lme4)
library(margins)
library(reprex)
library(plm)
library(devtools)
library(lfe)
library(Statamarkdown)
library(readxl)
library(starpolishr)
library(AER)
library(readxl)
library(naniar)
library(mice)
library(gtsummary)
library(VIM)
```

```{r data_imports, include = FALSE}

set.seed(100)

# import both datasets

rp <- read_dta("racial_polarization_winners.dta")
fin_seg<- read_dta("fin_seg.dta")

```

```{r data_cleaning_rp, include = FALSE}

# create a factor for year for the rp dataset
rp$year.f <- as.factor(rp$year)

# apply the condition that winner == 1 as in the Stata code for the rp regressions

rp_1<- rp %>%
  filter(winner == 1) %>%
  mutate(medincinterp = medincinterp/1000)

```

# Introduction

There is a large degree of variation in public goods spending across local governments. As a result, many scholars have worked to determine what factors may lead to the underprovision of public goods spending. Research in the past has associated racial diversity or changes in levels of diversity with the under-provision of public goods [@baqir_public_1999; @hopkins_diversity_2009].However, @trounstine_segregation_2016 argues that it is racial segregation, not diversity in and of itself, that results in the under-provision of public goods. Trounstine’s analysis consists of two main parts. First, she uses election and demographic data from 25 of America’s largest cities between 1990 and 2010 to run a multilevel mixed-effects linear regression with fixed effects for region and year and with random effects for cities in order to show that polarization increases with segregation. Trounstine measures segregation with Theil's H index, which measures the degree to which the diversity of a neighborhood differs from the diversity of the entire city. The second main part of Trounstine’s analysis looks at the ability of the Theil’s H segregation index to explain a variety of types of public expenditures by city using a sample of 2,637 cities with 13,742 city-year observations. Using linear regressions with fixed effects for cities and robust standard errors clustered by city, Trounstine finds a significant, negative correlation between segregation and public goods spending that is robust to the inclusion of a variety of relevant controls and an alternative specification in which the number of waterways is used as an instrument for segregation.

In this analysis, I first work to replicate the main results of @trounstine_segregation_2016 using R statistical software [@r]. The original data and Stata code made publically available by the author were downloaded via the Harvard Dataverse [@DVN/4LZXTY_2015]. I also make all of my code and analysis available on Github.^[ Link to my [Github repository](https://github.com/mburzillo/new_final_project) for this project.] I was successfully able to replicate the main results of @trounstine_segregation_2016 in R with the exception of some of the marginal analyses, which I was nevertheless able to replicate in Stata.

```{r data_prep_imputations, include = FALSE}

# select the relevant rows from the rp dataset

rp_impute <- rp %>%
  select("winner", "medincinterp",
             "biggestsplit", "H_citytract_multi_i", "diversityinterp",
             "pctasianpopinterp", "pctblkpopinterp", "pctlatinopopinterp",
             "pctrentersinterp", "pctcollegegradinterp", "biracial", "nonpartisan",
             "primary", "logpop", "year", "south", "midwest", "west", "geo_id2",
             "H_citytract_NHW_i", "whiteideology_fill2") %>%
  filter(winner == 1) %>%
  mutate(medincinterp = medincinterp/1000,
         year.f = as.factor(year))

# apply the conditions in Stata regression to the fin_seg dataset

fin_imp <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), dgepercap_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

# select the relevant variables to be imputed

fin_imp <- fin_imp %>%
  select(H_citytract_NHW_i, dgepercap_cpi, diversityinterp, pctblkpopinterp,
         pctasianpopinterp,pctlatinopopinterp, medinc_cpi, pctlocalgovworker_100,
         pctrentersinterp, pctover65, pctcollegegradinterp, logpop, geo_id2,
         chng5pctblk, chng5pctlatino, chng5pctasian, ideology_fill)

```

```{r obs_in_original_fin_and_cities, include = FALSE}
# basic filter for table for all regressions

# xtreg dgepercap_cpi H_citytract_NHW_i diversityinterp pctblkpopinterp
# pctasianpopinterp pctlatinopopinterp medinc_cpi pctlocalgovworker_100
# pctrentersinterp pctover65 pctcollegegradinterp logpop if totaltracts>1 &
# dgepercap_cpi~=0,fe vce(cluster geo_id2)

fin <-fin_seg %>%
  filter(totaltractsf >1 | is.na(totaltractsf)) %>%
  filter(!(is.na(dgepercap_cpi)), !(is.na(H_citytract_NHW_i)), !(is.na(diversityinterp)),
         !(is.na(pctblkpopinterp)), !(is.na(pctasianpopinterp)), !(is.na(pctlatinopopinterp)),
         !(is.na(medinc_cpi)), !(is.na(pctlocalgovworker_100)), !(is.na(pctrentersinterp)),
         !(is.na(pctover65)), !(is.na(pctcollegegradinterp)), !(is.na(logpop)))

# create dge variable used in regression

fin_dge <-fin %>% filter(dgepercap_cpi != 0)

# cities in original Trounstine

cities_trounstine_fp <- fin_dge %>%
  select(geo_id2) %>%
  unique() %>%
  pull()

# When we subset for only the cities in the original analysis, this about doubles the sample size still...

fin_imp_cities <- fin_imp %>%
  filter(geo_id2 %in% cities_trounstine_fp)

```

```{r obs_count, include = FALSE}

# calculate number of original observations

obs_fin_seg <- nrow(fin_imp)
obs_rp <- nrow(rp_impute)

```

One concern with the original analysis in @trounstine_segregation_2016 is the large amount of missing data, which substantially constrains the sample size used in the regression analysis. For example, the regression analysis in the main specification using the the original racial polarization dataset exludes %`r round((obs_rp - 91)/obs_rp, 2)` of the observations, and the regression analysis in the main specification using the original financial segregation data excludes %`r round((obs_fin_seg - 13742)/ obs_fin_seg, 2)` of observations from the original dataset. As an extension of @trounstine_segregation_2016, I impute missing values in the original data using the mice package in R, which generates multivariate imputations using chained equations [@mice]. Then, I use the multiply imputed datasets to re-estimate the original models, pooling the results to generate final pooled regression coefficients and parameters.

Comparing the results of the original regressions and those done with the imputed data yields similar big picture results in terms of the direction of the signs on the coefficients on the segregation indices. Like in @trounstine_segregation_2016, I find that segregation is positively associated with political polarization and negatively associated with spending on public goods. However, the magnitude of the effects in most specifications has diminished and most results become statistically insignificant. While these findings do not necessarily challenge the results of @trounstine_segregation_2016, they do call into question the relative importance of segregation in determining public goods spending and political polarization and suggest that the results of @trounstine_segregation_2016 may not be quite as robust as once thought.

# Literature Review

Despite some progress made towards racial equity in the U.S. on other fronts, residential racial segregation continues to be prevasive and deeply entrenched in society [@fischer_distinguishing_2004; @oliver_paradoxes_2010; @massey_american_1993]. Research suggests that this kind of segregation has political consequences, as political cleavages in segregated cities tend to have racial as well as spatial dimensions [@massey_american_1993]. Neighborhoods are often important actors within local politics because local governments provide many functions that are allocational in nature and concern geographical space [@trounstine_segregation_2016]. Thus, when neighborhoods are divided on racial lines as well as spacial lines, it is natural to expect higher degrees of racial polarization as a result.

Studying residential segregation is difficult because its effects tend to differ by geographic level. On the neighborhood level, the kind of geographic racial isolation brought on by residential segregation has been associated with racial intolerance, resentment, and competition between racial groups [@oliver_paradoxes_2010]. Living within segregated neighborhoods has also been associated with holding negative stereotypes and perceptions about out groups [@eric_oliver_intergroup_2003]. As a result, homogeneous neighborhoods have been associated with increased racial tension and political polarization in comparison to integrated, diverse neighborhoods. However, at the city or metropolitan level, the opposite seems to be true: when considering larger geographic areas, diversity and integration are correlated with racial tension, competition, prejudice, lower levels of cooperation, and lower spending on public goods [@oliver_paradoxes_2010; @baqir_public_1999; @hopkins_diversity_2009]. While these differences in the expected effect of segregation on the geographic level may seem confusing at first, they make sense as they suggest that the most severely segregated areas are those that are diverse overall, but have many homogeneous neighborhoods. Thus, while people of different races co-exist within a highly segregated city, they live separately within their own neighborhoods, which creates an environment ripe for racial tension [@trounstine_segregation_2016]. It is thus not simply the level of diversity or integration that matters for racial harmony and cooperation, but their patterns within a larger geographic framework [@trounstine_segregation_2016; @oliver_paradoxes_2010; @bharathi_public_2018].

Political polarization along racial lines may lead to decreased public spending and goods provision because groups may have different preferences, which can make compromise hard, and because groups may preceive a disutility in out-groups receiving public goods expenditure [@baqir_public_1999]. @einstein_divided_2012 found evidence that racial segregation predicts large political divisions at the metropolitan level and that these divisions can create a lack of willingness to compromise and collaborate on local policy problems. @trounstine_segregation_2016 finds similar results at the city level: that residential racial segregation is associated with both increased political division and decreased public spending. Thus, these authors suggest that it is the combination of homogenous neighborhoods within a much larger, diverse geographic area that leads to increased political polarization and reduced public goods spending in local governments.

More recently, some scholars have called this hypothesis and its importance into question. For example, @lee_ethnic_2018 finds evidence that larger inequalities within the political system favoring socially powerful groups, not local diversity patterns leading to decreased cooperation, may be a better explanation of failures in public goods provision in diverse areas. Others suggest that additional factors, such as income segregation, may be important confounding factors in public goods provision. @an_its_2018, for example, suggests that the more closely related income inequality is to racial inequality, the less investment is made in public goods, and that this interaction was a better predictor of public goods spending patterns than measures of diversity. There is, in fact, a variety of evidence suggesting that it is meaningful to consider the effects of income inequaltiy and diversity and segregation jointly [@an_its_2018; @massey_american_1993]. Given the active debate in the literature over the relaitonships between diversity, segregation, public spending, and other factors, it is increasingly important to re-examine previously reported findings as a means of robustness checks.

# Replication

I was able to successfully replicate all of the main results from Trounstine (2016). All regressions and tables were fully replicated in R. As an example, in table X, I replicate Table 1 from page 713 of @trounstine_segregation_2016, which is also included as Figure 1 for comparison. However, I was unable to successfully replicate Trounstine (2016)’s marginal effects analyses and margins plots using R. There does not yet appear to be a built-in R function to calculate marginal or predicted effects or to generate margins plots from the complicated multi-level models employed in the original paper, and creating such a function was outside of the scope of this analysis. Nevertheless, these results were successfully replicated in Stata.

There was one interesting outcome from my attempt to replicate the original Stata code in R. Due to the differences in R and Stata in dealing with missing values, the results of my first replication of analyses using the financial segregation dataset were slightly different than the results in Stata from the original paper. In order to exclude cities from her analysis with only one census tract, Trounstine conditions her regressions in Stata such that the value for the number of census tracts is greater than one. In Stata, this does not remove missing values, whereas in R, it does. Since this variable is used only as a conditional filter and not as a regression variable (and thus, the observations with missing values for number of census tracts are not dropped), Trounstine’s analysis includes 14 cities and 58 observations with missing values for census tracts in addition to cities with two or more census tracts. This is a potential oversight on the part of the author, and I would suggest also dropping observations with missing census tract data, or else imputing them given that the aim of her condition was to exclude cities without a sufficient number of census tracts. Nevertheless, dropping these values did not have much of an effect on the subsequent results, likely because they represent a small proportion of the overall sample.

# Extension

One concern with the original analysis in @trounstine_segregation_2016 is the large amount of missing data values in her original datasets. Because both R and Stata drop any observations with missing values for any of the variables used in a regression, this can exlcude a large portion of the data from the analysis and potentially bias the results if the data is not missing completely at random. A large amount of data is missing in both of the main datasets used in the analyses of @trounstine_segregation_2016. The main model specification using the racial polarization dataset exludes `r (obs_rp - 91)` of the original  `r obs_rp` observations, or  %`r round(100*(obs_rp - 91)/obs_rp,2)` of the data. Additionally, the main model specification using the financial segregation data set excludes `r round(obs_fin_seg - 13742,2)` of the original `r obs_fin_seg` observations, or %`r round(100*(obs_fin_seg - 13742)/ obs_fin_seg,2)` of observations from the original dataset.

In an attempt to better deal with the problem of missing data in @trounstine_segregation_2016, I impute missing values in the original data using the mice package in R, which generates multivariate imputations using chained equations [@mice]. While there are a variety of different imputation methods that could have been employed, multiple imputation (such as the multiple multivariate imputations generated by mice) is desirable because instead of inputting a single value such as the mean for missing values, it instead uses the distribution of the available data to estimate multiple potential values for the missing data. As a result, multiple imputation helps to account for the uncertainty inherent in the imputation process and allows for the calculation of standard errors around estimators. Thus, multiple imputation allows the researcher to more accurately assess the of uncertainty in the analysis in general.

Before performing the multiple imputations on the datasets, I first examined the missing data for patterns. To better understand any potential patterns in missing data, I plotted the pattern of missingness and created a histogram showing the frequency of missing values for those variables with missing values in figure X1 for the racial polarization data and figure X2 for the financial segregation data. Looking at figure X1, the histogram shows that the variables for the largest vote split along racial lines and the variable indicating whether or not the election had candidates of more than one race had by far the largest percentages of missing data, missing %`r round(sum(is.na(rp_impute$biggestsplit))/nrow(rp_impute)*100,2)` and  %`r round(sum(is.na(rp_impute$biracial))/nrow(rp_impute)*100,2)`, respectively. It is important to note that the variable for the largest vote split along racial lines is the dependent variable in the regresison analysis, and thus, we are missing a large percentage of this key variable. From the plot on the right, we can see that approximately 42% of observations are complete (have no missing values). There seem to be a correspondence between missing a value for the largest vote split and missing a value for the biracial election indicator. There also seem to be about 2% of values for wich most of the variables are missing. However, most observations are not missing more than 2-3 values.

# Appendix

```{r include = FALSE}

## regression 1 Table 1

# perform a multi-level mixed-effects linear regression of biggest split on the multi-group H index with controls for Herfindahl Diversity Index, % Asian interpolated, % Black interpolated, % Latino interpolated, median household income interpolated, % renters interpolated, percent college grads interpolated, a dummy indicating if there were biracial candidates in the election, a dummy  indicating if the election was nonpartisan, an indicator if the election was a primary, and the log population. The regression also includes fixed effects for year and region and random effects for cities (geo_id2). Diversity is  included because of the claim that more diverse populations will have more heterogeneous political preferences. The other fixed effect control variables are included because they are known to be potentially correlated with segregation. winner == 1 requires that there was a winner in the election.

# regression specification. Apply REML = FALSE. The default for lmer() in R is
# to optimize the restricted (residual) maximum likelihood (REML) whereas in
# xtmixed in Stata the default is to optimize the log likelihood (which you can
# do in lmer() in R by setting REML = FALSE), so this will match the STATA code

m1 <- lmer(biggestsplit ~ H_citytract_multi_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), data = rp_1, REML=FALSE)

## regression 2 Table 1

# Perform the same regression as for Column 1 above but this time regress on the two-group calculation of Theil's H interpolated. It is unclear where the "diversity" variable is coming from, as the only variable related to diversity in the specificed dataset is diversityinterp. This is a problem that will need to be resolved. Essentially, this regression just tests an alternate specification to the first in which the segregation index in defined in terms of two-groups rather than multiple.

m2 <- lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), data = rp_1, REML=FALSE)

# perform the same regression as for column 2, but now add in a measure of
# average white political ideology. The point of this is to control for ideology
# so that the author can separate any effects of segregation from any potential
# effects of white political ideology. This is important because it is possible
# that segregation is just a proxy for white conservatism and thus failure to
# include this variable could lead to ommitted variable bias.

m3 <- lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + whiteideology_fill2 + year.f + south + midwest + west + (1 | geo_id2), data = rp_1, REML=FALSE)

```

```{r results="asis", echo = FALSE, fig.cap= "THIS IS A CAPTION IS IT WORKING WHAT IS GOING ON!"}

# create a stargazer table of the results

t1 <- stargazer(m1, m2, m3, omit = c("year", "south", "midwest", "west"),
                header = FALSE,
                title = "\\textbf{Racial Polarization in Segregated Cities}",
                single.row = TRUE,
                no.space = FALSE,
                keep.stat = c("n"),
                covariate.labels = c("Multigroup H Index", "White/Nonwhite H Index", "Diversity", "Percent Asian", "Percent Black", "Percent Latino", "Medain HH Income (1000s)", "Percent Renters", "Percent College Degree", "Biracial Contest", "Nonpartisan Election", "Primary Election", "Population (logged)", "White Ideology", "Constant"),
                dep.var.labels = c("Racial Divide"),
                column.labels = c("Multigroup H Index", "White/Nonwhite H Index", "Ideology Controls"),
                notes = "This table displays the results of the replication\neffort of the analyses conducted in Trounstine (2016) using the racial polarization dataset.\nThe replication code can be found on my personal Github repository, which is linked in the introduction. The original table is displayed for comparison as Figure 1.)",
                notes.align = "l")

# note: "This table displays the results of the replication effort of the analyses conducted in Trounstine (2016) using the racial polarization dataset. The replication code can be found on my personal [Github repository](https://github.com/mburzillo/new_final_project). The original table is displayed for comparison as Figure 1.")

```

![This is the original table displaying the results of the models for the effects of residential segregation on political polarization from page 713 of @trounstine_segregation_2016. My replication of these results can be found in Table 1 in this paper.](table_1.png)

```{r felm, include = FALSE}

# Main Analysis 2

## regression 1 Table 2

# The following is a fixed effects model regression with clustering around
# geo_id2. It mandates that total census tracts in the city is greater than 1,
# which we want because our measure of evenness of racial spread is constant for
# cities with only one tract by our definition (which requires comparing tracts
# within cities). It also mandates that Direct General Expenditures (DGE) per
# capita, CPI adjusted is not equal to 0, which would be problematic and
# indicate a potential data error. This regression regresses the DGE per capita
# on the the two-group calculation of  Theil's H interpolated. Controls are
# included for diversity and the % populations of Blacks, Asians, and Latinos to
# help identify whether it is actually diversity that drives down spending.These
# demographic controls also help to account for the fact that whites and
# minorities tend to perfer different levels of government spending in general.
# This will help isolate the effects of segregation v. other forms of diversity
# on public spending. Other controls for Median household income, CPI adjusted,
# percent over 65, percent college grads, percent local government workers, log
# population help control for additional factors related to segregation and
# expenditure. Controls for percent renters interpolated and Median household
# income also help to control for the general wealth of cities, which could be a
# potential driver of public spending. Fixed effects for cities are also
# included  so that the author can examine the effect of segregation in the same
# city over  time, which also helps control for many other factors not otherwise
# taken into account (such as city age)

# apply the conditions in Stata regression

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), dgepercap_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

# felm(y~x | factor(fe1) + factor(fe2) | 0 | clustervar)

felm1 <- felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

## regression 2 Table 2

# This regression is the same as above except instead of using diversity, a new
# control is added for the five year changes in racial group shares. The goal
# here is to see whether or not changes in diversity are a driving factor
# (rather than absolute levels of diversity).

felm2 <- felm(dgepercap_cpi ~ H_citytract_NHW_i + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + chng5pctblk + chng5pctlatino + chng5pctasian+ medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

## Regression 3 Table 2

# Column 3

# As in column 3 of table 1, this regression simply adds in a control for  mean  ideology of city residents from General Social Survey (GSS) to control for the  possibility that ideology that segregated cities are more ideologically  conservative and that the conservative nature of segregated cities is what actually drives the effect.

felm3 <- felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop + ideology_fill | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

```

```{r main_fp_tables, results = "asis", echo = FALSE}

# create main table for table 3

table2 <- stargazer(felm1, felm2, felm3, omit = c("year", "south", "midwest", "west"), 
                    header = FALSE,
                    title = "\textbf{Effect of Segregation on Overall per Capita City Expenditures}",
                    covariate.labels = c("White/Nonwhite H Index", "Diversity", "Percent Black", "Percent Asian", "Percent Latino", "Change Percent Black", "Change Percent Latino", "Change Percent Asian", "Medain HH Income (1000s)", "Percent Government Worker", "Percent Renters", "Percent 65+", "Percent College Degree", "Population (logged)", "White Ideology", "Constant"))

```

```{r, include = FALSE}

# Main Analysis 3

## regression 1 Table 3

# Column 1

# The following is a fixed effects model regression with clustering once again around geo_id2. Again, the if statements at the end mandate that total census tracts in the city is greater than 1, which we want because our measure of evenness of racial spread is constant for cities with only one tract by our definition (which requires comparing tracts within cities). They also mandate that the highways per capita, CPI adjsuted with no cap extend (lagged 5 years) is not equal to 0, which would be problematic and indicate a potential data error. The dependent variable here is highways per capita, CPI adjsuted with no cap extend (lagged 5 years) and it is regressed on the two group calculation of Theil's H, interpolated. Controls are added for diversity, percent population of Blacks, Asians, and Latinos, and median household CPI adjusted income as well as for % local government worker hundreds, percent rentership, percent over 65, percent college graduates, and the log of the population. This essentially examines the effect of segregation on the specific public good of highways. As in the previous table's regressions, fixed effects for cities are also included so that the author can examine the effect of segregation in the same city over time, which also helps control for many other factors not otherwise taken into account (such as city age) *

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), highwayspercapNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_1 <- felm(highwayspercapNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

## regression 2 Table 3

# Column 2

# This is the same regression as column 1, but this time the dependent variable
# is police per capita, CPI adjusted with no cap extend. This essentially
# examines the effect of segregation on the specific public good of police per
# capita.

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), policepercapNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_2 <- felm( policepercapNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# Column 3
fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), parkspercapNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

# This is again the same regression, but this time the dependent variable is
# parks per capita, CPI adjusted with no cap extend. This essentially examines
# the effect of segregation on the specific public good of parks.

felm2_3 <- felm(parkspercapNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# Column 4

# This is again the same regression, but this time the dependent variable is
# sewers per capita, CPI adjusted with no cap extend. This essentially examines
# the effect of segregation on the specific public good of sewers.

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), sewerspercapNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_4 <- felm(sewerspercapNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# Column 5

# This is again the same regression, but this time the dependent variable is
# Welfare, Health, Housing per capita, CPI adjusted, no cap expend. This
# essentially examines the effect of segregation on the specific public good of
# welfare and housing.

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), welfhoushealthNC_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_5 <- felm(welfhoushealthNC_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

# Column 6

# This is again the same regression, but this time the dependent variable is own
# source of general revenue per capita, CPI adjusted. This essentially examines
# the effect of segregation on the specific public good of own source revenue.

fin_seg_sub <- fin_seg %>%
  filter(totaltractsf > 1 | is.na(totaltractsf), genrevownpercap_cpi != 0) %>%
  mutate(medinc_cpi = medinc_cpi/1000)

felm2_6 <- felm(genrevownpercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2, data = fin_seg_sub)

```

```{r iv_regs, include = FALSE}

## regression 1 Table 5

# Top Row

# Column 1

# This is an instrumental variable regression with DGE per capita, adjusted by the CPI, as the dependent variable and total number of waterways as the instrumental variable for segregation. The instrument includes the log of the population in the first stage because the number of waterways is positively correlated to population and population is positively correlated with segregation. The control variables are the same as those in the regressions for Table 2 except for 2 changes: since the number of waterways does not change, there is no need for city fixed effects. Instead, we include fixed effects for the region and year. A lagged version of the dependent variable (dgepercap_cpilag) is added to the regressions to account for the high correlation between  observations for the same city over time and because changes in local budgets typically are incremental (NOTE: why exactly do we do this/why does it help/why not do it in the non-intstrumental regressions?).

fin_seg_1 <- fin_seg %>%
  filter(dgepercap_cpi != 0) %>%
  mutate(medincinterp = medincinterp/1000)

glimpse(fin_seg_1$medincinterp)

# good explanation: http://eclr.humanities.manchester.ac.uk/index.php/IV_in_R

# iv = ivreg(second stage regression | instrument1 + instrument2 )

ivreg1 <- ivreg(dgepercap_cpi ~ H_citytract_NHW_i + dgepercap_cpilag + diversityinterp +
                  pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +
                  pctcollegegradinterp + northeast + south + midwest + y5 + 
                  y6 + y7 + y8 + y9 | total_rivs_all + 
                  logpop + dgepercap_cpilag + diversityinterp + pctblkpopinterp + 
                  pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +   
                  pctcollegegradinterp + northeast + south + 
                  midwest + y5 + y6 + y7 + y8 + y9, data = fin_seg_1)

# Column 2

# this is the same regression as for column 1 but now we use the dependent variable highwayspercapNC_cpi to assess the effect of the instrumental  variable on highways per capita *

fin_seg_2 <- fin_seg %>%
  filter(highwayspercapNC_cpi != 0) %>%
  mutate(medincinterp = medincinterp/1000)

ivreg2 <- ivreg(highwayspercapNC_cpi ~ H_citytract_NHW_i + highwayspercapNC_cpilag + diversityinterp +
                  pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +
                  pctcollegegradinterp + northeast + south + midwest + y5 + 
                  y6 + y7 + y8 + y9 | total_rivs_all + 
                  logpop + highwayspercapNC_cpilag + diversityinterp + pctblkpopinterp + 
                  pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +   
                  pctcollegegradinterp + northeast + south + 
                  midwest + y5 + y6 + y7 + y8 + y9, data = fin_seg_2)

# Column 3

# this is the same regression as for column 1 but now we use the dependent variable policepercapNC_cpi to assess the effect of the instrumental variable on police per capita

fin_seg_3 <- fin_seg %>%
  filter(policepercapNC_cpi != 0) %>%
  mutate(medincinterp = medincinterp/1000)

ivreg3 <- ivreg(policepercapNC_cpi ~ H_citytract_NHW_i + policepercapNC_cpilag + diversityinterp +
                  pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +
                  pctcollegegradinterp + northeast + south + midwest + y5 + 
                  y6 + y7 + y8 + y9 | total_rivs_all + 
                  logpop + policepercapNC_cpilag + diversityinterp + pctblkpopinterp + 
                  pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +   
                  pctcollegegradinterp + northeast + south + 
                  midwest + y5 + y6 + y7 + y8 + y9, data = fin_seg_3)

# Column 4 this is the same regression as for column 1 but now we use the
# dependent variable parkspercapNC_cpi to assess the effect of the instrumental
# variable on parks per capita

fin_seg_4 <- fin_seg %>%
  filter(parkspercapNC_cpi != 0) %>%
  mutate(medincinterp = medincinterp/1000)

ivreg4 <- ivreg(parkspercapNC_cpi ~ H_citytract_NHW_i + parkspercapNC_cpilag + diversityinterp +
                  pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +
                  pctcollegegradinterp + northeast + south + midwest + y5 + 
                  y6 + y7 + y8 + y9 | total_rivs_all + 
                  logpop + parkspercapNC_cpilag + diversityinterp + pctblkpopinterp + 
                  pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +   
                  pctcollegegradinterp + northeast + south + 
                  midwest + y5 + y6 + y7 + y8 + y9, data = fin_seg_4)

# Bottom Row

# Column 1

# this is the same regression as for column 1 but now we use the dependent
# variable sewerspercapNC_cpi to assess the effect of the instrumental variable
# on sewers per capita

fin_seg_5 <- fin_seg %>%
  filter(sewerspercapNC_cpi != 0) %>%
  mutate(medincinterp = medincinterp/1000)

ivreg5 <- ivreg(sewerspercapNC_cpi ~ H_citytract_NHW_i + sewerspercapNC_cpilag + diversityinterp +
                  pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +
                  pctcollegegradinterp + northeast + south + midwest + y5 + 
                  y6 + y7 + y8 + y9 | total_rivs_all + 
                  logpop + sewerspercapNC_cpilag + diversityinterp + pctblkpopinterp + 
                  pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +   
                  pctcollegegradinterp + northeast + south + 
                  midwest + y5 + y6 + y7 + y8 + y9, data = fin_seg_5)

# Column 3

# this is the same regression as for column 1 but now we use the dependent variable genrevownpercap_cpi to assess the effect of the instrumental variable on own source revenue per capita

fin_seg_6 <- fin_seg %>%
  filter(genrevownpercap_cpi != 0) %>%
  mutate(medincinterp = medincinterp/1000)

ivreg6 <- ivreg(genrevownpercap_cpi ~ H_citytract_NHW_i + genrevownpercap_cpilag + diversityinterp +
                  pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +
                  pctcollegegradinterp + northeast + south + midwest + y5 + 
                  y6 + y7 + y8 + y9 | total_rivs_all + 
                  logpop + genrevownpercap_cpilag + diversityinterp + pctblkpopinterp + 
                  pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +   
                  pctcollegegradinterp + northeast + south + 
                  midwest + y5 + y6 + y7 + y8 + y9, data = fin_seg_6)

# Column 2

# this is the same regression as for column 1 but now we use the dependent
# variable welfhoushealthNC_cpi to assess the effect of the instrumental
# variable on welfare per capita

fin_seg_7 <- fin_seg %>%
  filter(welfhoushealthNC_cpi != 0) %>%
  mutate(medincinterp = medincinterp/1000)

ivreg7 <- ivreg(welfhoushealthNC_cpi ~ H_citytract_NHW_i + welfhoushealthNC_cpilag  + diversityinterp +
                  pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +
                  pctcollegegradinterp + northeast + south + midwest + y5 + 
                  y6 + y7 + y8 + y9 | total_rivs_all + 
                  logpop + welfhoushealthNC_cpilag + diversityinterp + pctblkpopinterp + 
                  pctasianpopinterp + pctlatinopopinterp + medincinterp +
                  pctlocalgovworker_100 + pctrentersinterp + pctover65 +   
                  pctcollegegradinterp + northeast + south + 
                  midwest + y5 + y6 + y7 + y8 + y9, data = fin_seg_7)

summary(ivreg7)

```

# Bibliography