### 2020-04-30 stargazer_reprex.Rmd

---
title: "Title"
author: "Author"
output: 
  pdf_document:
    fig_caption: yes
    extra_dependencies: ["rotating"]
header_includes:
  \usepackage{float}
  \usepackage{dcolumn}
  \usepackage{rotating}
  \restylefloat{table}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.position = "!H")
library(tidyverse)
library(stargazer)
library(reprex)
```

```{r results = "asis", echo = FALSE}

library(tidyverse)
library(stargazer)

output <- lm(mpg ~ cyl + disp + wt, data = mtcars)
output2 <- lm(mpg ~ cyl + disp + wt + drat, data = mtcars)

t1 <- stargazer(output, output2, 
                header = FALSE,
                title = "\\textbf{Practice Table}",
                notes = "Notes Could Go Here???????????? This is an example. Notes Could Go Here???????????? This is an example. Notes Could Go Here???????????? This is an example. Notes Could Go Here???????????? This is an example. ",
                notes.align = "l",
                float = FALSE,
                table.placement = "H")

```